---
title: "Detterent Effects of SQFs on Arrests in New York City: A Spatial Analysis for 2006"
author:
  - Jonas Frost
institute: 
  - jonas.frost@studserv.uni-leipzig.de \linebreak Master's Student at Institute for Sociology \linebreak Departement of Social Sciences and Philosophy \linebreak University of Leipzig \normalsize
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output:
    beamer_presentation:
      theme: "Frankfurt"
      color: "dolphin"
      slide_level: 2
bibliography: ../Literatur.bib 
csl: ../zfs.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, fig.height = 5)
knitr::opts_knit$set(root.dir = "~/_UNI/_MA/3_Semester/Spatial_Econometrics/AA_Project_SpatEcon")
```

```{r data, cache=FALSE, include=FALSE}
source("Code/Makefile.R")
theme_set(theme_bw())
```

# Background

## Motivation

  * NYPD claims Stop, Question and Frisks (SQFs) deter crime
  * SQFs are highly controversial because of racial profiling and police violence allegations
  * In 2013 federal court ruled SQF tactics in New York had to be changed $\rightarrow$ numbers drastically decreased
  * Question remains: Did SQFs deter crime? 

## The current study's features

  * Spatial analyses of all New York SQFs and arrests in 2006
  * Micro-econometric design at the level of individual SQFs 
  * Computationally intensive approach 
  * Trying to close in on casual inference

# Theory

## Considerations

 * SQFs are based on **broken windows hypothesis** [@Fagan.2000; @Wilson.1982]
 * SQFs might change criminal's assesments of probability of being caught
 * Effect could be detering or displacing
 * Crime is clustered at specific locations

## Theoretic Model

\small

  * $U(c \mid \ell_i)$: Utility of crime at location $i$
  * $C(c \mid \ell_i)$: Cost of crime at location $i$
  * $p_{\ell_i}$: Probability of being caught at location $i$
  * $U(\neg c \mid \ell_i)$: Utility of not choosing crime at location $i$
  * $C(\neg c \mid \ell_i)$: Cost of not choosing crime at location $i$
  * $(N_{sqf} \mid \ell_i, t)$: Number of SQFs at location $i$ and time $t$
  * $(N_c \mid \ell_i, t)$: Number of crimes at location $i$ and time $t$

\normalsize

### Decision Model
$$
U(c \mid \ell_i) - C(c \mid \ell_i) \cdot p_{\ell_i} \  > \ U(\neg c \mid \ell_i) - C(\neg c \mid \ell_i)
$$
$$
p_{\ell_i} \ = \ f((N_{sqf} \mid \ell_i, t), \ (N_c \mid \ell_i, t), \ \boldsymbol{\cdot} \ )
$$

## Hypothesis

### Deterrence Hypothesis

An SQF leads to a lesser number of arrests in a contained area of space and time.

# Data

## Data

  * Data with geo-coordinates of SQFs and arrests in 2006 aquired from \color{blue} [New York City's open data platform](https://opendata.cityofnewyork.us/) \color{black}
  
```{r density_map}
nyc_shape_owin <-
    nyc_zip_shape_ll |>
    st_transform(crs = 6345) |>
    st_simplify(dTolerance = 25) |>
    as.data.frame()

density_arrests <- density.ppp(arrests.ppp[[1]])
density_sqf <- density.ppp(sqf.ppp[[1]])


gg1 <-
  as.data.frame(density_arrests) |>
  ggplot(mapping = aes(x = x,
                       y = y,
                       fill = value)) +
  geom_tile() +
  theme_void() +
  labs(title = "Kernel Density:\nNew York Arrests 2006") +
  scale_fill_viridis(name = "Density",
                     limits = range(c(density_arrests$v, density_sqf$v), na.rm = TRUE))


gg2 <- as.data.frame(density_sqf) |>
  ggplot(mapping = aes(x = x,
                       y = y,
                       fill = value)) +
  geom_tile() +
  theme_void() +
  labs(title = "Kernel Density:\nNew York SQFs 2006") +
  scale_fill_viridis(name = "Density",
                     limits = range(c(density_arrests$v, density_sqf$v), na.rm = TRUE))

plot_grid(
    gg1,
    gg2,
    ncol = 2
    ) +
  theme(plot.margin = margin(1, 1, 1, 1, "cm"),
        plot.background = element_rect(color = "black"))

```

## Data

```{r time, message=FALSE}
tmp1 <-
  data_arrest |>
  filter(ARREST_DATE < "2007-01-01",
         ARREST_DATE >= "2006-01-01") |>
  count(ARREST_DATE) |>
  mutate(from = "arrests") |>
  rename(date = ARREST_DATE)

tmp2 <-
  data_sqf |>
  filter(datestop < "2007-01-01",
         datestop >= "2006-01-01") |>
  count(datestop) |>
  mutate(from = "SQFs") |>
  rename(date = datestop)

tmp <- bind_rows(tmp1, tmp2)

more_arrests <-
  tmp |>
  group_by(date) |>
  filter(n == max(n)) |>
  ungroup() |>
  count(from) 
```

  * `r nrow(filter(data_final, is_sqf == 1))` SQFs and `r nrow(filter(data_arrest, ARREST_DATE < "2007-01-01", ARREST_DATE >= "2006-01-01"))` arrests were included in the analyses
  * only `r more_arrests$n[1]` days in 2006 with more arrests than SQFs
  * Amounts of SQFs and arrests follow a temporal periodicity $\rightarrow$ number of arrests might decrease after SQF because of this periodicity only

```{r time_plot, message=FALSE}
ggplot(tmp,
       aes(x = date,
           y = n,
           color = from)) +
  geom_smooth(linewidth = 0.8) +
  geom_line() +
  ylab("amount") +
  scale_color_manual(name = "", 
                     values = c("SQFs" = "brown3",
                                "arrests" = "cornflowerblue")) +
  labs(title = "number of SQFs and arrests over time")
```

# Methods

## Methods

  * Ripley's Cross K-function
    + are arrests spatially (anti-)clustered around SQFs?
  * Micro-econometric analysis at level of individual SQFs
    + count amount of arrests in buffer around each SQF
    + varying time from -10 to +10 days from the event
    + varying buffer radius of 1km, 5km and 10km
    + poisson point process with 500 points per day as random control events
    + redo arrest counts for random control
    
## Micro-econometric Design


```{r map_buffer, warning=FALSE, fig.height=6}
points_sqf_ll <- 
  data_final |>
  filter(datestop == "2006-03-10") |>
  # sample_n(800) |>
  mutate(across(.cols = ends_with("coord"),.fns = as.numeric),
         geometry = st_multipoint(as.matrix(cbind(xcoord, ycoord))),
         is_sqf = factor(is_sqf,
                         levels = c("1", "0"),
                         labels = c("SQF", "Control"))
  ) |>
  # select(geometry) |>
  st_as_sf(coords = c("xcoord","ycoord"), crs = st_crs("WGS84"))

points_arrests_ll <-
  data_arrest |>
  filter(ARREST_DATE == "2006-03-10") |>
  sample_n(800) |>
  select(Longitude, Latitude) |>
  na.omit() |>
  as.matrix() |>
  st_multipoint() |>
  st_sfc(crs = st_crs("WGS84"))

manhattan_ll <- 
  nyc_shape_ll |>
  filter(BoroName == "Manhattan")

gg1 <-
  ggplot() +
  geom_sf(data = nyc_shape_ll, fill = "white") +
  geom_sf(data = points_sqf_ll,
          aes(color = as.factor(is_sqf))) +
  labs(color = "") +
  theme_void() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        legend.position = "bottom") +
  labs(title = "(1) SQFs and Control Events on 2006-03-10")

this_sqf <- points_sqf_ll[57,]
buffer <- buffer_sqf <- st_buffer(this_sqf, dist = 1000)

gg2 <-
  ggplot() +
  geom_sf(data = manhattan_ll, fill = "white") +
  geom_sf(data = buffer, color = "red", fill = "white") +
  geom_sf(data = this_sqf, color = "red") +
  theme_void() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
  labs(title = "(2) 1km Buffer around each Event")

int_2006 <- st_intersection(buffer, points_arrests_ll)

gg3 <-
  ggplot() +
  geom_sf(data = manhattan_ll, fill = "white") +
  geom_sf(data = buffer, color = "red", fill = "white") +
  geom_sf(data = int_2006, alpha = 0.5) +
  geom_sf(data = this_sqf, color = "red") +
  theme_void() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
  labs(title = "(3) Count number of Arrests in Buffer")

gg1

```

## Micro-econometric Design

```{r manhattan_plot, cache=FALSE, fig.height=6}
plot_grid(
  gg2,
  gg3
)
```

# Results

## Cross K-function

  * "The expected number of *arrests* lying within a distance r of a typical *SQF*, standardised by dividing by the intensity of *arrests*" [@Baddeley.2016: 594]

```{r k_cross, fig.height=3}
K1 <- Kcross[[1]]

ggplot(mapping = aes(x = r)) +
    geom_line(mapping = aes(y = K1$iso, color = "K-Cross")) +
    geom_line(mapping = aes(y = K1$theo, color = "CSR"), linetype = "dashed") +
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
    xlab("Distance in meters") +
    ylab("K-hat") +
    scale_color_manual(values = c("K-Cross" = "cornflowerblue", "CSR" = "black"), name = "") +
    theme_bw() +
    labs(title = paste0("Cross K-function for Arrests and SQFs in 2006")) +
    theme(legend.position = c(0.3, 0.8), legend.title = element_blank())
```

  * Arrests are highly clustered around SQFs
  * Problem: this approach **doesn't consider time**
  * Possible interpretation: there are crime hot spots, where arrests and SQFs are clustered together

## Micro Approach

  * More arrests around SQFs then around random points
  * No change in number of arrests around random control events
  * Number of arrests before and after SQFs are barely different

```{r box_plots, cache=TRUE}
ylim1 <- boxplot.stats(filter(data_final_long, dist == 1) |>
                         pull(arrests))$stats[c(1, 5)]
ylim5 <- boxplot.stats(filter(data_final_long, dist == 5) |>
                         pull(arrests))$stats[c(1, 5)]
ylim10 <- boxplot.stats(filter(data_final_long, dist == 10) |>
                          pull(arrests))$stats[c(1, 5)]

data_final_long |>
  filter(time_to_sqf >= -10,
         time_to_sqf <= 10,
         dist == 1 & arrests <= ylim1[2] |
           dist == 5 & arrests <= ylim5[2] |
           dist == 10 & arrests <= ylim10[2]) |>
  mutate(dist = as.factor(dist) |>
           recode("1" = "1km",
                  "5" = "5km",
                  "10" = "10km")) |>
  ggplot(
    mapping = aes(
      x = factor(time_to_sqf, levels = as.character(c(-10:10))),
      fill = is_sqf,
      y = arrests
    )) +
  geom_boxplot(notch = TRUE, outlier.shape = NA) +
  geom_vline(xintercept = "0", color = "brown3") +
  scale_x_discrete(breaks = as.character(c(-10:10)), drop = FALSE) +
  labs(title = "Number of arrests in radius around SQF and random points") +
  guides(shape = "none") +
  xlab("Days to/from SQF") +
  ylab("Number of Arrests") +
  scale_fill_manual(name = "",
                    values = c("0" = "cornflowerblue",
                               "1" = "brown3"),
                    labels = c("CSR", "SQF")) +
  theme_bw() +
  facet_grid(vars(dist), scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(subtitle = "Outliers removed for better display.")

```

## Micro Approach

```{r regression_tab, message=FALSE, warning=FALSE, include=FALSE}
tab <-
  stargazer(
    lmer_fit_1,
    lmer_fit_5,
    lmer_fit_10,
    lmer_fit_all,
    header = FALSE,
    digit.separator = " ",
    intercept.top = TRUE,
    intercept.bottom = FALSE,
    no.space = FALSE,
    column.sep.width = "3pt",
    omit.stat = "ll",
    font.size = "tiny",
    dep.var.labels = "Number of Arrests",
    covariate.labels = c("Intercept", "Time to Event", "Other SQFs", "After Event", "SQF or Control", "SQF*After"),
    notes.append = TRUE,
    notes = c("All models are fitted using OLS fixed-effects for the 5 boroughs of New York.", "Model with all radii of observation includes additional fixed-efffects for radii.", "Full dataset was sampled to include 50000 events."),
    notes.align = "l",
    column.labels = c("in 1km radius", "in 5km radius", "in 10km radius", "in all radii"),
    model.names = FALSE,
    model.numbers = FALSE
  )

tab <- sub('^.+\\caption.+$','', tab)
tab <- sub('^.+\\label.+$','', tab)
```

```{r regression_tab2, cache=FALSE, results='asis'}
cat(tab, sep = "\n")
```

## Conclusion

  * Deterring effects of SQFs on arrests are not stable throughout all observed spatial windows
  * Found effects stay small


# Discussion

## Forthcomings

  * Other authors [@Weisburd.2015; @Rosenfeld.2017] find similarly small deterring effects of SQFs 
  * Attempt to estimate effects of individual SQFs instead of looking at effects at fixed spatial windows
  * Control for temporal and spatial variance
  * Randomly generated control group

## Limitations

  * Effects between arrest rates and SQF rates cannot be fully entangled
  * Arrest rates only a vague measurement of crime $\rightarrow$ many other unconsidered factors influence arrest rates
  
\vspace*{1cm}

### Conclusion
Though the current investigation is limited in its interpretation, the NYPDs and other proponents claim that SQF practices effectively deter crime cannot be supported.

## References

\small
